---
title: "Analysis of Weekday Effects"
output: html_notebook
---

# Load support functions and libraries
```{r}

# Load libraries
library("httr")
library("jsonlite")
library("quantmod")
library("mgcv")

source("./support_functions.R")

# Read the saved data
btc.data <- read.csv("./BTC_historical_march_10.csv")
eth.data <- read.csv("./ETH_historical_march_10.csv")

# Create some support functions
btc.data <- as.data.frame(btc.data)

getDate <- function(x){
  date.stamp <- as.Date(strsplit(as.character(x)," ")[[1]][1], origin="1970-01-01")
  return(date.stamp)
}

getHour <- function(x){
  hour <- as.numeric(strsplit(strsplit(as.character(x)," ")[[1]][2],":")[[1]][1])
  return(hour)
}

getWeekday <- function(x){
  weekday <- weekdays(as.Date(as.character(x), origin="1970-01-01"))
  return(weekday)
}

```

#  Analysis for BTC
```{r}
btc.data <- as.data.frame(btc.data)
btc.data$date.stamp <- as.Date(unlist(lapply(as.character(btc.data$Time.Stamp), getDate)), origin="1970-01-01")
btc.data$hour <- unlist(lapply(btc.data$`Time.Stamp`, getHour))
btc.data$weekday <- unlist(lapply(btc.data$`Time.Stamp`, getWeekday))

plot(as.numeric(as.character(btc.data$Close)), type="lines")

btc.data$price.detrended <- as.numeric(as.character(btc.data$Close))/probabilisticSmoothing(as.numeric(as.character(btc.data$Close)), trials = 2000, lag.u = 0, lag.sd = 24) - 1
btc.data$weekday <- as.factor(btc.data$weekday)

summary(lm(price.detrended ~ weekday, data=btc.data))
```
The plot above shows the detrended time series, which allows us to control for trend in the calculation of weekday effects.
The summary table shows the estimated efect of each weekday on the overall volume change (see the Estimate column). Saturday appears to have the strongest positive effect, with Friday the weakest. 

# Now aggregate onto the daily level and check the effect
```{r}

btc.data.daily <- subset(btc.data, btc.data$hour == "12")
plot(as.numeric(as.character(btc.data.daily$Close)), type="lines")

btc.data.daily$price.detrended <- as.numeric(as.character(btc.data.daily$Close))/probabilisticSmoothing(as.numeric(as.character(btc.data.daily$Close)), trials = 2000, lag.u = 0, lag.sd = 4) - 1
btc.data.daily$weekday <- as.factor(btc.data.daily$weekday)

# Plot the detrended value
plot(btc.data.daily$price.detrended, type="lines")

# Plot the Weekday vs the Price
plot(btc.data.daily$weekday, btc.data.daily$price.detrended)

# Regression of day vs Price
lm.fit <- lm(price.detrended ~ weekday, data=btc.data.daily)

```
The plot above shows the estimated effect of each weekday, after aggregating. As expected Saturday, Sunday and Monday appear to have the strongest effect. The effects are not strongly significant to a scientific level, but there is likely a positive effect based on the sample. This can be seen in the table below:
```{r}
summary(lm.fit)
```

#The following analysis will check for combination of day and time effects
```{r}
# Now look at the day/time combination
btc.data$weekday.hour <- paste(btc.data$weekday,"_",btc.data$hour, sep="")

first.value <- btc.data[1,'weekday.hour']
counter <- 1
btc.data$weekday.hour.num <- NA
for (i in 1:nrow(btc.data)){
  if (btc.data[i,'weekday.hour'] == first.value){
    counter <- 1
    btc.data[i,'weekday.hour.num'] <- counter
  } else {
    btc.data[i,'weekday.hour.num'] <- counter
  }
  counter <- counter + 1
}

btc.data$weekday.hour.num <- as.numeric(as.factor(btc.data$weekday.hour))

cyclical.vars <- cyclicalTransformation(btc.data$weekday.hour.num, 168)
btc.data <- cbind(btc.data, cyclical.vars)

# Linear model fit

lm.fit <- lm(price.detrended ~ x + y, data=btc.data)
plot(btc.data$price.detrended, type="lines")

# Non linear fit

gam.fit <- mgcv::gam(price.detrended ~ s(x, k=4) + s(y, k=4), data=btc.data, family="gaussian")
{plot(btc.data$price.detrended, type="lines")
lines(predict(gam.fit), col="red")}

{plot(btc.data$weekday.hour.num, btc.data$price.detrended)
points(btc.data$weekday.hour.num, predict(gam.fit), col="red")}


```
The plot above shows the estimated fit based on the time of day and weekday combination. You can see that there is a lot of noise, and as such there is unlikely to be a strong pattern based on time of day. However we can check this as we receive more data.

# Analysis of ETH
```{r}


eth.data <- as.data.frame(eth.data)
eth.data$date.stamp <- as.Date(unlist(lapply(eth.data$`Time.Stamp`, getDate)), origin="1970-01-01")
eth.data$hour <- unlist(lapply(eth.data$`Time.Stamp`, getHour))
eth.data$weekday <- unlist(lapply(eth.data$`Time.Stamp`, getWeekday))

plot(as.numeric(as.character(eth.data$Close)), type="lines")

eth.data$price.detrended <- as.numeric(as.character(eth.data$Close))/probabilisticSmoothing(as.numeric(as.character(eth.data$Close)), trials = 2000, lag.u = 0, lag.sd = 24) - 1
eth.data$weekday <- as.factor(eth.data$weekday)

summary(lm(price.detrended ~ weekday, data=eth.data))
```
# Aggregate onto a daily level
```{r}
eth.data.daily <- subset(eth.data, eth.data$hour == "12")
plot(as.numeric(as.character(eth.data.daily$Close)), type="lines")

eth.data.daily$price.detrended <- as.numeric(as.character(eth.data.daily$Close))/probabilisticSmoothing(as.numeric(as.character(eth.data.daily$Close)), trials = 2000, lag.u = 0, lag.sd = 4) - 1
eth.data.daily$weekday <- as.factor(eth.data.daily$weekday)

# Plot the detrended value
plot(eth.data.daily$price.detrended, type="lines")

# Plot the Weekday vs the Price
plot(eth.data.daily$weekday, eth.data.daily$price.detrended)

# Regression of day vs Price
lm.fit <- lm(price.detrended ~ weekday, data=eth.data.daily)


# Result - no significant evidence of day of the week affecting the price, however there 
#is a slight hint that saturdays tend to have higher prices
```
The plot above shows the estimated effect of weekday on the ETH price. The pattern is a little inconsisten in this plot.
```{r}
summary(lm.fit)
```
The table above shows the results of the linear regression for ETH against the weekday. The summary confirms that monday and saturday have the highest effect. However there is not a strong difference to the other days. 
```{r}
# Now look at the day/time combination
eth.data$weekday.hour <- paste(eth.data$weekday,"_",eth.data$hour, sep="")

first.value <- eth.data[1,'weekday.hour']
counter <- 1
eth.data$weekday.hour.num <- NA
for (i in 1:nrow(eth.data)){
  if (eth.data[i,'weekday.hour'] == first.value){
    counter <- 1
    eth.data[i,'weekday.hour.num'] <- counter
  } else {
    eth.data[i,'weekday.hour.num'] <- counter
  }
  counter <- counter + 1
}

cyclical.vars <- cyclicalTransformation(eth.data$weekday.hour.num, 168)
eth.data <- cbind(eth.data, cyclical.vars)


# Linear model fit
lm.fit <- lm(price.detrended ~ x + y, data=eth.data)
{plot(eth.data$price.detrended, type="lines")
lines(predict(lm.fit), col="red")}

# Non linear fit
gam.fit <- mgcv::gam(price.detrended ~ s(x, k=4) + s(y, k=4), data=eth.data, family="gaussian")
{plot(eth.data$price.detrended, type="lines")
lines(predict(gam.fit), col="red")}

{plot(eth.data$weekday.hour.num, eth.data$price.detrended)
points(eth.data$weekday.hour.num, predict(gam.fit), col="red")}




```
The image above shows a similar pattern to the BTC coin price, but again there is no strong evidence of time effects.


# Summary 

 In summary, both linear models suggest an increase in price on the weekends. While not statistically significant to a 5% level p vaue,
 it would definitey be worth keeping an eye on and validating on future data. 

 Finally with regard to hours in the day, there appears to be a very slight trend heading into Saturday.
 This could be due to noise in the current sample and I would hesitate to draw strong conclusions given the current small sample size.


# Strategy backtest
Buy on Wednesday at 4pm and sell on Saturday at 4pm
```{r}
# Test the buy Wednesday Sell Saturday strategy on BTC
money <- 1
owned.coins <- 0
for (i in 1:nrow(btc.data)){
  if (btc.data[i,'weekday.hour'] == "Wednesday_16"){
    owned.coins <- money/as.numeric(as.character(btc.data[i,'Open']))
  }
  
  if (btc.data[i,'weekday.hour'] == "Saturday_16"){
    money <- owned.coins*as.numeric(as.character(btc.data[i,'Open']))
    print(money)
    owned.coins <- 0
  }
}


```
End with a 12% increase, compared to general BTC losses on the overall market.
```{r}
money <- 1
owned.coins <- 0
for (i in 1:nrow(eth.data)){
  if (eth.data[i,'weekday.hour'] == "Wednesday_16"){
    owned.coins <- money/as.numeric(as.character(eth.data[i,'Open']))
  }
  
  if (eth.data[i,'weekday.hour'] == "Saturday_16"){
    money <- owned.coins*as.numeric(as.character(eth.data[i,'Open']))
    print(money)
    owned.coins <- 0
  }
}
```

On the ETH test, the strategy would yield a 38% increase over the tested timerange. Compared to a 3% increase over the entire timerange.
```{r}
eth.data$Close[nrow(eth.data)]/eth.data$Close[1]
```
In summary, this methodology performed better than the market over the given timerange of December to March.
